### 2024-07-28 paper/aamas/aamas25-submission.tex
tmp

%%%% For anonymized submission, use this
%\documentclass[sigconf,anonymous]{aamas} 
%%%% For camera-ready, use this
\documentclass[sigconf]{aamas}

\usepackage{balance} % for balancing columns on the final page

\setcopyright{ifaamas}
\acmConference[AAMAS '25]{Proc.\@ of the 24th International Conference
on Autonomous Agents and Multiagent Systems (AAMAS 2025)}{May 19 -- 23, 2025}
{Detroit, Michigan, USA}{A.~El~Fallah~Seghrouchni, Y.~Vorobeychik, S.~Das, A.~Nowe (eds.)}
\copyrightyear{2025}
\acmYear{2025}
\acmDOI{}
\acmPrice{}
\acmISBN{}

%%% Use this command to specify your EasyChair submission number.
%%% In anonymous mode, it will be printed on the first page.
\acmSubmissionID{<<EasyChair submission id>>}

\title{Title}

%%% Provide names, affiliations, and email addresses for all authors.

\author{Francesco Bertolotti}
\affiliation{
  \institution{Università degli Studi di Milano}
  \city{Milan}
  \country{Italy}}
\email{francesco.bertolotti@di.unimi.it}

\author{Walter Cazzola}
\affiliation{
  \institution{Università degli Studi di Milano}
  \city{Milan}
  \country{Italy}}
\email{walter.cazzola@di.unimi.it}

\author{Gianluca Aguzzi}
\affiliation{
  \institution{Università di Bologna}
  \city{Bologna}
  \country{Italy}}
\email{gianluca.aguzzi@unibo.it}

\author{Mirko Viroli}
\affiliation{
  \institution{Università di Bologna}
  \city{Bologna}
  \country{Italy}}
\email{mirko.viroli@unibo.it}

\keywords{Multi-Agent, Differetiable Environment, World Model}

\begin{document}

\pagestyle{fancy}
\fancyhead{}
\input{sects/abstract.tex}
\maketitle 
\input{sects/acks.tex}

\bibliographystyle{ACM-Reference-Format} 
\bibliography{ml+nn,foundations,strings,aosd,my_work,grammars,software_engineering,reflection,dsu,dsl,pl,splc,programming,roles,oolanguages,testing,./bibs/local}

\end{document}