### 2025-01-08 paper/icml25/first/sects/related.tex
minor + related (WIP)

\section{Related Work}

\paragraph{Differentiable Engines}
Game engines~\cite{Bellemare13,Balla23} and Physics engines \cite{Todorov12,Coumans16} have been wildely used in the field of policy learning. Recently, there has been a trend towards differentiable engines, which allows for the use of gradient-based optimization methods, such as backpropagation. Notable examples are \cite{Freeman21,Howell22,Genesis24}. These framework may be built on top of auto differentiation libraries such \cite{Bradbury18,Ansel24} (this is the case of VMAS) or directly with analytical gradients \cite{Carpentier18,Werling21}. Further, several works tackle the challenge of the non-smoothness of the contact dynamics \cite{Degrave19,Moritz20}

\paragraph{Deep Reinforcement Learning}
Deep reinforcement learning has revolutionized policy learning tasks by combining reinforcement learning with deep neural networks. Foundational works like DQN \cite{Mnih13} and AlphaZero \cite{Silver17} demonstrated human-level performance in complex environments such as Atari games and chess.

\paragraph{Multi-Agent Reinforcement Learning}

\paragraph{Reinforcement Learning on Differentiable Engines}