### 2025-01-21 paper/icml25/first/appendices/arch.tex

\section{Architectures}\label{apx:arch}

\begin{table}[t]
        \centering
        \input{tables/shacwm.tex}
        \caption{SHAC/\fname{}/\fnamer{} Hyperparameters}
        \label{apx:tab:shac}
\end{table}

\begin{table}[t]
        \centering
        \input{tables/ppo.tex}
        \caption{PPO Hyperparmeters}
        \label{apx:tab:ppo}
\end{table}