### 2025-01-21 paper/icml25/first/appendices/additional.tex
minors + appendix (WIP)

\section{Additional Experiments}

\begin{figure}[t]
    \centering
    \includegraphics[width=\columnwidth]{figs/main-mlp.pdf}
    \caption{Comparison between \fname{}, PPO, and SHAC for increasing number of agents for Dispersion, Transport, Discovery, and Sampling scenarios with the MLP architecture.
    We show the mean and standard deviation of the normalized rewards across $3$ runs. For the Transformer architecture see \Cref{fig:experiments}.
    }
    \label{apx:fig:experiments-mlp}
\end{figure}

\begin{figure*}[!t]
    \centering
    \includegraphics[width=\textwidth]{figs/ablation-transformer.pdf}
    \caption{Comparison between \fname{} with and without transition network, labelled \fnamer{}, for increasing number of agents for Dispersion, Transport, Discovery, and Sampling scenarios.}
    \label{fig:ablation}
\end{figure*}

\begin{figure}[!t]
    \centering
    \includegraphics[width=\columnwidth]{figs/ablation-mlp.pdf}
    \caption{Comparison between \fname{} with and without transition network, labelled \fnamer{}, for increasing number of agents for Dispersion, Transport, Discovery, and Sampling scenarios.}
    \label{fig:ablation}
\end{figure}