### 2025-04-22 paper/ecai2025/sects/conclusion.tex
update to ecai 2025 template

\section{Conclusions \& Future Works}\label{sect:conclusion}
In this paper, we proposed \fname{}, an extension of SHAC, to provide a reinforcement learning solution that is both sample efficient, able to work in both single-agent and multi-agent scenarios, and capable of handling non-differentiable scenarios. We demonstrated that this solution maintains alignment with SHAC (\RQ{1}), outperforms the current state-of-the-art PPO in both single-agent and multi-agent scenarios (\RQ{2}), and exhibits promising scaling capabilities in multi-agent settings with emergent collaborative behaviors (\RQ{3}).

Despite these encouraging results, several directions for future work remain. First, we aim to explore scalability beyond the current limitation of tens of agents. Additionally, as evidenced in the discovery scenario, the current approach struggles when agents have highly partial observations of the environment. Furthermore, we plan to evaluate our approach in other multi-agent environments, such as those presented from the cooperative AI community~\cite{DBLP:journals/corr/abs-2211-13746} and also in adversarial settings.