### 2025-07-26 paper/camera/sects/acknowledgements.tex

\section*{Acknowledgements}

\begin{itemize}
    \item Computational resources provided by INDACO Platform, which is a project of High Performance Computing at the University of MILAN http://www.unimi.it
    \item This work was partly supported by the MUR project “T-LADIES” (PRIN2020TL3X8X).
    \item TODO (unibo)
\end{itemize}