### 2025-08-04 paper/camera/sects/limitations.tex

\section{Limitations}

While the proposed method shows promise, it is not a one-size-fits-all solution. Environments vary significantly in complexity and dynamics, and several limitations must be acknowledged.

\paragraph{World Model Learning and Cold-Start Problem}
A central challenge lies in learning the world model, which depends on rich agent-environment interactions. Sparse or delayed rewards create a cold-start problem: until the world model is sufficiently trained, agents struggle to explore effectively. This can be addressed by seeding with a cold-start dataset---manually designed or generated using algorithms like PPO or MPPI~\cite{Alvarez24}---to bootstrap the world model and provide a stable training foundation.

\paragraph{Baseline Performance and Implementation}
Despite careful attention to implementation, including reference to best practices (e.g., the well-known ``37 Implementation Details of Proximal Policy Optimization'' blog~\cite{Shengyi22}), we observed that PPO and MAPPO performed suboptimally in our experiments. To ensure the validity of these findings, we cross-verified results using a third-party implementation~\cite{Bettini24}, which yielded consistent outcomes.

\paragraph{Policy Collapse and Robustness}
Although early stopping is primarily used to conserve computational resources, it can give the false impression that the algorithm is inherently robust to policy collapse. In reality, we observed multiple instances of partial policy collapse---some of which recovered---and a few cases of complete collapse without recovery. One example of partial collapse appears in the five-agent variant of the dispersion scenario shown in \Cref{apx:fig:experiments-mlp}.

\paragraph{Partial Observability Challenges}
The discovery scenario poses considerable difficulty for all algorithms due to inherent partial observability. Agents rely solely on simulated LiDAR for perception, providing only a narrow, local view. This makes it difficult for any method---including \fname{}---to construct accurate world models, often resulting in suboptimal performance and suggesting that partial observability remains challenging for the current approach.

\paragraph{Computational Overhead}
\fname{} addresses SHAC's limitations by adding transition and reward networks, but this increases computational overhead. Training burden is especially pronounced with many agents, where the transition network must capture complex inter-agent dynamics. While this trade-off is often acceptable, it significantly slows training and demands substantial computational resources.

\paragraph{Interpretability}
Regarding interpretability, \fname{} does not offer significant advantages over its predecessors. As the policy is governed by a neural network, understanding or explaining its behavior remains a challenging task. This limitation is inherent to most deep reinforcement learning methods and highlights the ongoing need for techniques that enhance policy transparency and interpretability.