### 2025-08-05 paper/camera/sects/acknowledgements.tex
ready to submit

\section*{Acknowledgements}

\begin{itemize*}
    \item Computational resources provided by INDACO Platform, which is a project of High Performance Computing at the University of MILAN http://www.unimi.it
    \item This work was partly supported by the MUR project “T-LADIES” (PRIN2020TL3X8X), by the Italian PRIN project ``CommonWears'' (2020 HCWWLP), and by the Gemma 2 Google Research Award 2024.
    %\item TODO (unibo)
\end{itemize*}