### 2025-08-17 paper/camera/sects/acknowledgements.tex
separated appendix from the main paper

\section*{Acknowledgements}

\begin{itemize}
    \item Computational resources provided by INDACO Platform, which is a project of High Performance Computing at the University of MILAN http://www.unimi.it
    \item This work was partly supported by the MUR project ``T-LADIES'' (PRIN2020TL3X8X).
    \item This work was partly supported by the Italian PRIN project ``CommonWears'' (2020 HCWWLP).
    \item Computational cloud resources were provided by the Gemma 2 Google Research Award 2024.
    %\item TODO (unibo)
\end{itemize}